Covariate analysis of CMC RNASeq data reprocessed with Star-Feature counts pipeline (ACC and DLPFC with TMM normlaisation and Custom Design 1)

### 2016-11-06 code/Rmd/CMC_ACC_And_DLPFC_StarAligned_TMM_CD1.Rmd
Added markdown to process DLPFC and ACC together with a random effects model with a custom desing with all important variables

---
title: "Covariate analysis of CMC RNASeq data reprocessed with Star-Feature counts pipeline (ACC and DLPFC with TMM normlaisation)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knitr)
library(knit2synapse)
library(synapseClient)

@@

knit2synapse::knitToFolderEntity(file = "./CMC_ACC_And_DLPFC_StarAligned_TMM.Rmd",
                                 parentId ="syn7501880",
                                 entityName = 'ACC and DLPFC - STAR Reprocessed - with TMM')
```

@@

thisFileName <- 'CMC_ACC_And_DLPFC_StarAligned_TMM.Rmd'

@@

### Normalisation (iterative design)
Since many covariates are correlated, re-normalising COUNTS with an iterative design matrix. Here "Dx.Tissue" is chosen as the primary variable of interest (i.e., covariate selection is conditioned on diagnosis and tissue)

Adding Instituion and Gender a priori to variable selection as they are correlated with Dx.Tissue
```{r iterativeNorm, results='asis'}
# Primary variable of interest
postAdjustCovars = c('Dx.Tissue', 'Institution', 'Gender');

# Assign residual covariates
residualSigCovars = preAdjustedSigCovars
covariatesEffects = residualSigCovars$Effects.significantCovars[setdiff(residualSigCovars$significantCovars, postAdjustCovars)]
postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects))) %>% unique()

loopCount = 0 
while(length(residualSigCovars$significantCovars)!=0 && loopCount <= 100){
  writeLines(paste('Using following covariates in the model:',
                   paste(postAdjustCovars, collapse=', '),
                   'as fixed effects'))

@@

  # Find PC of residual gene expression and significant covariates that are highly correlated with PCs
  residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                                COVARIATES[, residCovars, drop=F], 
                                                'adjusted design(voom-normalized)',
                                                isKeyPlot=TRUE)

  # Add postadjusted covariates (if any)
  covariatesEffects = residualSigCovars$Effects.significantCovars[setdiff(residualSigCovars$significantCovars, 
                                                                          c(postAdjustCovars))]

  postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects)))
  loopCount = loopCount + 1
}

modelStr <- paste(paste(gsub('_','\\\\_',postAdjustCovars), collapse=', '),
                  'as fixed effects')

tmp <- paste('Using following covariates in the final model:', modelStr)
```
`r tmp`

@@

tmp1 = DE %>%
  dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.2)) %>%
  dplyr::select(ensembl_gene_id, Comparison, Tissue) %>%
  group_by(Comparison, Tisue) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id)))

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "ACC and DLPFC - STAR Reprocessed - with TMM", parentId = parentId)
CODE <- synStore(CODE)

>>>>

---
title: "Covariate analysis of CMC RNASeq data reprocessed with Star-Feature counts pipeline (ACC and DLPFC with TMM normlaisation and Custom Design 1)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knitr)
library(knit2synapse)
library(synapseClient)

@@

knit2synapse::knitToFolderEntity(file = "./CMC_ACC_And_DLPFC_StarAligned_TMM_CD1.Rmd",
                                 parentId ="syn7501880",
                                 entityName = 'ACC and DLPFC - STAR Reprocessed - with TMM - Custom Design 1')
```

@@

thisFileName <- 'CMC_ACC_And_DLPFC_StarAligned_TMM_CD1.Rmd'

@@

### Normalisation (custom design)
Normalising COUNTS with a custom design matrix. Here Instituiton, Gender, LibraryBatch, Age\\_of\\_Death, PMI, RIN, RIN2, EV.1, EV.2, EV.3, EV.4, EV.5 are adjusted conditioned on Dx.Tissue with Individual\\_ID as random effects.
```{r iterativeNorm, results='asis'}
# Primary variable of interest
postAdjustCovars = c('Dx.Tissue', 'Institution', 'Gender', 'LibraryBatch', 'Age\\_of\\_Death',
                     'PMI', 'RIN', 'RIN2', 'EV.1', 'EV.2', 'EV.3', 'EV.4', 'EV.5');

writeLines(paste('Using following covariates in the model:',
                 paste(postAdjustCovars, collapse=', '),
                 'as fixed effects and Individual_ID as random effect'))

@@

# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(RESIDUAL.GENE_EXPRESSION, 
                                              COVARIATES[, residCovars, drop=F], 
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)
```

@@

tmp1 = DE %>%
  dplyr::filter(adj.P.Val <= 0.05, abs(logFC) >= log2(1.2)) %>%
  dplyr::select(ensembl_gene_id, Comparison, Tissue) %>%
  group_by(Comparison, Tissue) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id)))

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "ACC and DLPFC - STAR Reprocessed - with TMM - Custom Design 1", parentId = parentId)
CODE <- synStore(CODE)
